Searching for circuit complexity lower bounds using SAT solvers

### 2022-05-07 report/draft1.tex
report progress, times for extension variable experiments

Finding lower bounds on the size of Boolean circuits computing given Boolean functions is one of the most fundamental open problems in computer science. Despite the fact that the majority of functions have exponential circuits, the best lower bounds proven on unrestricted circuits are only linear~\cite{boppana} and we still do not have optimal circuits or close upper and lower bounds for many important functions. Results in this area can be used to separate computational complexity classes, possibly even P and NP (i.e. to show that a large class of fundamental and practically useful problems do not have polynomially fast algorithms).~\cite{arora}

If functions with circuits of a given complexity could be identified by an efficient algorithm, this could be used to construct efficient learning algorithms for circuits of a smaller complexity.~\cite{carmosino} Furthermore, such an algorithm could be used to break pseudorandom generators (i.e. distinguish random from pseudorandom inputs), which would have major implications in the field of cryptography.~\cite{arora}

@@

One approach to automating the search for efficient circuits is to reduce the problem of designing a correct circuit (logical design synthesis) to the Boolean satisfiability problem (SAT). Existing algorithms (SAT solvers) can then be used to solve the problem, either finding a correct circuit of fixed size or generating a proof that none exists~\cite{kulikov}.

@@

Circuits are a popular model of non-uniform computation. In the past there have been theoretical proofs of lower bounds for restricted circuit classes, e.g. AC0\cite{furst81}\cite{ajtai83}, AC0[k] (i.e. AC0 additionally with MOD-k gates)\cite{razborov87}\cite{smolensky87}, and monotone circuits\cite{razborov85}. However, no lower bounds better than 5n-c\cite{iwama02} have been found for general circuits. Furthermore, there are results stating that it is impossible to use circuit lower bounds to separate P and NP by relativisation\cite{baker75}, algebrization\cite{aaronson09}, or natural proofs\cite{razborov94}. It has been proved that circuit lower bounds are hard to prove using resolution and resolution with k-DNFS, and it is conjectured that they are also hard under stronger proof systems such as Frege systems\cite{raz}. Due to this difficulty in theoretical progress, we would like to investigate the problem empirically.

Kamath et al~\cite{kamath} propose a reduction from logical design synthesis to SAT which fixes the DNF structure of the circuit and number of disjunctions used. Experiments using this reduction to find minimal circuits for 2-4 bit adders and multipliers are reported in~\cite{estrada}. Kojevnikov et al~\cite{kulikov} demonstrate a more general reduction allowing any circuit with a fixed number of gates, proving linear bounds for \(MOD^n_3\) circuits over different Boolean bases. This was done by checking for circuit 'building blocks' with 5 to 11 gates. In~\cite{kulikovlocal} the same reduction was used to improve large circuit building blocks by searching for smaller versions of their sub-circuits. Specifically, linear upper bounds were proved or verified for \(SUM\), \(MOD^n_4\), and \(MOD^n_3\) by checking for circuits with up to 13 gates.

@@

From these experiments we note that available solvers running on a personal computer can decide the existence of some circuits of size 11 in just a few hours, while others of size 11 and up can take many hours on a high-powered computing cluster. That is to say, the amount of time taken to solve comparable problems can vary considerably.

@@

We used three SAT solvers, all based on conflict-driven clause learning. This is an algorithm closely related to DPLL, and which also makes use of the resolution proof system.\cite{cdcl}\cite{krajicek}

MiniSAT\cite{minisat} was the winner of the 2005 SAT Competition and was last updated to version 2.2 in 2010\cite{minisat2010}. Despite being outdated, it is well-documented and the other solvers we used were based on it, making it a good baseline for performance.

PicoSAT is a solver inspired by miniSAT 1.14, with a focus on improving low-level performance by using optimised data structures\cite{picosat}. It was submitted in the 2010 SAT-Race.

@@

*I'm not sure why we are using picoSAT - there is still time to re-run the results if we want to replace it.

@@

We encoded the circuit existence problem as a Boolean formula using three reductions.

@@

The first is due to Kojevnikov et al~\cite{kulikov} who present it together with bounds on the growth rate of the formula. The basis of the circuit can be specified as any subset of \(B_2\).

@@

The second is due to Razborov~\cite{raz}, assumes the function outputs only a single Boolean value, and requires a circuit over the basis \(C_2\). Note that it has a lower growth rate (in terms of number of clauses) than the other two reductions, suggesting it encodes the requirements more efficiently. Let $f$ be the target function with n inputs, and let $t$ be the number of gates required in the circuit.

@@

  \item $y_{av}$ $(a \in {0,1}^n, v \in [t])$ is the value taken by node $v$ on circuit input $a$. This gives $O(t2^n)$ variables.
  \item $y_{a\nu v}$ $(a \in {0,1}^n, \nu \in {1,2}, v \in [t])$ is the value of the $\nu$-th predecessor to $v$ on input $a$. This gives $O(t2^n)$ variables.
  \item $Fanin(v)$ is 0 if $v$ is a $\neg$-gate and 1 otherwise. This gives $O(t)$ variables.
  \item $Types(v)$ - when $Fanin(v)=1$, this is 0 if $v$ is a $\land$-gate and 1 if $v$ is a $\lor$-gate. This gives $O(t)$ variables.
  \item $InputType_\nu(v)$ is 0 if the $\nu$-th predecessor to $v$ is a constant or an input, 1 if it is another gate. This gives $O(t)$ variables.
  \item $InputType'_\nu(v)$ - when $InputType_{\nu}(v) = 0$, this is 0 if the $\nu$-th predecessor to $v$ is a constant, 1 if it is an input. This gives $O(t)$ variables.
  \item $InputType''_\nu(v)$ - when $InputType_{\nu}(v) = InputType'_{\nu}(v) = 0$, this is the value of the constant $\nu$-th predecessor to $v$. This gives $O(t)$ variables.
  \item $InputVar_\nu(v,i)$ $(i \in [n])$ - when $InputType_{\nu}(v) = 0, InputType'_{\nu}(v) = 1$, this is 1 iff the $\nu$-th predecessor to $v$ is the $i$-th input. This gives $O(tn)$ variables.
  \item $INPUTVAR_\nu(v,i)$ is equal to $\bigvee_{i'<i} InputVar_\nu(v,i')$, which will be used to bound the fan-in below. This gives $O(tn)$ variables.
  \item $InputNode_\nu(v,v')$ $(v' < v)$ - when $InputType_{\nu}(v) = 1$, this is 1 iff $\nu$-th predecessor to $v$ is the gate $v'$. This gives $O(t^2)$ variables.
  \item $INPUTNODE_\nu(v,v')$ is analogous to $INPUTVAR_\nu(v,i)$. This gives $O(t^2)$ variables.

@@

  \item Constant predecessors take the correct values, giving $O(t)$ clauses:

@@

  \item Gates have at most one input per input predecessor, giving $O(tn^2)$ clauses:

@@

  \item INPUTVAR variables take the correct value, giving $O(tn)$ clauses:

@@

  \item Gates have at least one input per input predecessor, giving $O(t)$ clauses:

@@

  \item Input predecessors take the correct values, giving $O(tn \cdot 2^n)$ clauses:

@@

  \item Gates take the correct value, giving $O(t2^n)$ clauses.

@@

The third reduction is straightforward. It also assumes the function outputs only a single Boolean value and requires a circuit over the basis \(C_2\).

@@

We chose to investigate the existence of circuits up to 9 gates, as these were reliably solved within less than 9 hours on the hardware described.

\subsection{Comparison of Boolean functions}

We investigated the relationship between the size of a circuit and the time taken for a SAT solver to confirm or deny its existence. We encoded a range of simple symmetric functions using the Kulikov encoding *define them more clearly:

\begin{enumerate}
  \item AND on 7 and 8-bit inputs
  \item Parity on 3- and 4-bit inputs
  \item \(MOD^0_3\) and \(MOD^0_4\) on 3- and 4-bit inputs
  \item Majority on 3 and 4-bit inputs
  \item Partial parity - Explain why partial functions are useful for encoding lower bounds on big input-lengths. But also need to find a way to get good data for them (ie. more than 1 second per problem...)

@@

\begin{figure}[!ht]
%  \includegraphics[width=\textwidth]{images/times/8and kulikov.png}
\end{figure}
The size of the minimal circuit is shown in red. It was found that the time needed tended to spike on or around the optimal circuit size. This may be because it is difficult to confirm or rule out circuits that come close to satisfying the requirements. A similar phenomenon was reported in~\cite{estrada}.

@@

\subsection{Comparison of SAT solvers and reductions}

We investigated whether there was a relationship between the reduction used to encode a Boolean function, the solver used, and the time taken by the solver.

\begin{figure}[!ht]
  \includegraphics[width=\textwidth]{images/totals.png}
  \label{fig:totals}
  \caption{Comparison of times taken by SAT solvers on different reductions}

@@

It can be seen that MapleSAT on the Kulikov encoding performs the best. Apart from this, the effect of the reduction and the solver seem to be independent, with MapleSAT consistently performing better than MiniSAT and PicoSAT, and the naive and Kulikov reductions performing better than the Razborov reduction. It is possible that the efficient encoding of the Razborov reduction is detrimental to the performance of the solver. However, these results are only applicable to small values of n. **however I should take out the constants and re-run the experiments to be sure it’s not that..

\subsection{Random functions}

16 truth tables were generated for 3 and 4-bit inputs using a pseudorandom number generator (specifically, xorshift128+ due to the Javascript implementation) simulating a uniform distribution. The tables were encoded with the Kulikov reduction and solved using MapleSAT. The times taken to check for circuits of different sizes are shown in
change to 64 and show the tables

*images*

The distribution of minimal circuit sizes needed to implement these functions are shown in \ref{fig:3rand_sizes} and \ref{fig:4rand_sizes}.

\begin{figure}[!ht]
  \includegraphics[width=\textwidth]{images/3rand_sizes.png}
  \label{fig:3rand_sizes}
  \caption{Sizes of optimal circuits for 3-bit functions with randomly chosen truth tables}
\end{figure}
\begin{figure}[!ht]
  \includegraphics[width=\textwidth]{images/4rand_sizes.png}
  \label{fig:4rand_sizes}
  \caption{Sizes of optimal circuits for 4-bit functions with randomly chosen truth tables}

A further 32 truth tables were generated for 4-bit inputs. While we expect a large proportion of circuits to be hard for large input sizes~\cite{arora}, the usual counting argument breaks down for such small values of n (ask for help - why?). However, using the SAT solver we can check the number of hard circuits directly.

*image*

@@

@@

@@

\subsection{Further directions}

>>>>

Finding lower bounds on the size of Boolean circuits computing given Boolean functions is one of the most fundamental and most challenging problems in computer science. Despite the fact that the majority of functions require exponential circuits, the best lower bounds proven on unrestricted circuits are only linear~\cite{boppana} and we still do not have optimal circuits or close upper and lower bounds for many important functions. Results in this area can be used to separate computational complexity classes, possibly even P and NP (i.e. to show that a large class of fundamental and practically useful problems do not have polynomially fast algorithms).~\cite{arora}

If functions with circuits of a given complexity could be identified by an efficient algorithm, this could be used to construct efficient learning algorithms for circuits of a smaller complexity; this would resolve one of the biggest questions in learning theory.~\cite{carmosino} Furthermore, such an algorithm could be used to break pseudorandom generators (i.e. distinguish random from pseudorandom inputs).~\cite{arora} As a result, this problem has strong connections in both learning theory and cryptography.

@@

One approach to automating the search for efficient circuits is to reduce the problem of designing a correct circuit (logical design synthesis) to the Boolean satisfiability problem (SAT). Existing algorithms (SAT solvers) can then be used to solve the problem, either finding a correct circuit of fixed size or generating a proof that none exists~\cite{kulikov}. SAT solvers have become increasingly powerful in recent years *cite* and now have many industrial applications*cite*; the wide range of heuristics implemented by different SAT solvers means that one suited to the problem at hand could be found or developed. *cite*

*how to extend it*

@@

Circuits are a popular model of non-uniform computation. In the past there have been theoretical proofs of lower bounds for restricted circuit classes, e.g. AC0\cite{furst81}\cite{ajtai83}, AC0[k] (i.e. AC0 additionally with MOD-k gates)\cite{razborov87}\cite{smolensky87}, and monotone circuits\cite{razborov85}. However, no lower bounds better than 5n-c\cite{iwama02} have been found for general circuits. Furthermore, there are results stating that it is impossible to use circuit lower bounds to separate P and NP by relativising methods\cite{baker75}, algebrizing methods\cite{aaronson09}, or (under standard crpytpgraphic assumptions) so-called 'natural proofs'\cite{razborov94}. It has been proved that circuit lower bounds are hard to prove using resolution and resolution with k-DNFS, and it is conjectured that they are also hard under stronger proof systems such as Frege systems\cite{raz}. Due to this difficulty in theoretical progress, we would like to investigate the problem empirically.

Kamath et al~\cite{kamath} propose a reduction from logical design synthesis to SAT which fixes the DNF structure of the circuit and number of disjunctions used. Experiments using this reduction to find minimal circuits for 2-4 bit adders and multipliers are reported in~\cite{estrada}. Kojevnikov et al~\cite{kulikov} demonstrate a more general reduction allowing any circuit with a fixed number of gates, proving linear bounds for \(MOD^n_3\) circuits over different Boolean bases. This was done by checking for circuit 'building blocks' with 5 to 11 gates. In~\cite{kulikovlocal} Kulikov et al. used the same reduction to improve large circuit building blocks by searching for smaller versions of their sub-circuits. Specifically, linear upper bounds were proved or verified for \(SUM\), \(MOD^n_4\), and \(MOD^n_3\) by checking for circuits with up to 13 gates (they remarked that checking for existence of a Boolean circuit of size 13 was out of reach).

@@

From these experiments we note that available solvers running on a personal computer can decide the existence of some circuits of size 11 in just a few hours, while others of size 11 and 12 can take many hours on a high-powered computing cluster (13 and above were not investigated). That is to say, the amount of time taken to solve comparable problems can vary considerably.

@@

We used three SAT solvers, all based on conflict-driven clause learning. This is an algorithm closely related to DPLL, and which also makes use of the resolution proof system.\cite{cdcl}\cite{krajicek} As mentioned earlier, this means that they cannot decide general circuit lower bounds with a less than exponential-length proof.\cite{raz} However, this asymptotic bound still allows for reasonably-sized proofs for small inputs such as the ones investigated in this paper.

MiniSAT\cite{minisat} was the winner of the 2005 SAT Competition and was last updated to version 2.2 in 2010\cite{minisat2010}. Despite being old, it is well-documented and the other solvers we used were based on it, making it a good baseline for performance. It was also one of the solvers used to decide circuit existence in \cite{kulikov}.

PicoSAT is a solver inspired by miniSAT 1.14, with a focus on improving low-level performance by using optimised data structures\cite{picosat}. It was submitted in the 2010 SAT-Race, and was also one of the solvers used in \cite{kulikov} and the main one in \cite{kulikovlocal}.

@@

Finally, for use on the Oxford ARC service *fill in the part about glucose-syrup aaaaaaaa*

@@

We encoded the circuit existence problem as a Boolean formula using three reductions. Let $t$ be the target function with $n$ inputs, and let $N$ be the number of gates required in the circuit.

@@

The first is due to Kojevnikov et al~\cite{kulikov} who present it together with bounds on the growth rate of the formula. In particular, the growth rate of the number of clauses is $O(N \cdot 2^n)$. The basis of the circuit can be specified as any subset of \(B_2\).

@@

The second is due to Razborov~\cite{raz}, assumes the function outputs only a single Boolean value, and requires a circuit over the basis \(C_2\). Note that it has a lower growth rate (in terms of number of clauses) than the other two reductions - $O(N \cdot 2^n)$ as opposed to $O(N^3 \cdot 2^n)$ - suggesting it encodes the requirements more efficiently.

@@

  \item $y_{av}$ $(a \in {0,1}^n, v \in [N])$ is the value taken by node $v$ on circuit input $a$. This gives $O(N \cdot2^n)$ variables.
  \item $y_{a\nu v}$ $(a \in {0,1}^n, \nu \in {1,2}, v \in [N])$ is the value of the $\nu$-th predecessor to $v$ on input $a$. This gives $O(N \cdot 2^n)$ variables.
  \item $Fanin(v)$ is 0 if $v$ is a $\neg$-gate and 1 otherwise. This gives $O(N)$ variables.
  \item $Types(v)$ - when $Fanin(v)=1$, this is 0 if $v$ is a $\land$-gate and 1 if $v$ is a $\lor$-gate. This gives $O(N)$ variables.
  \item $InputType_\nu(v)$ is 0 if the $\nu$-th predecessor to $v$ is a constant or an input, 1 if it is another gate. This gives $O(N)$ variables.
  \item $InputType'_\nu(v)$ - when $InputType_{\nu}(v) = 0$, this is 0 if the $\nu$-th predecessor to $v$ is a constant, 1 if it is an input. This gives $O(N)$ variables.
  \item $InputType''_\nu(v)$ - when $InputType_{\nu}(v) = InputType'_{\nu}(v) = 0$, this is the value of the constant $\nu$-th predecessor to $v$. This gives $O(N)$ variables.
  \item $InputVar_\nu(v,i)$ $(i \in [n])$ - when $InputType_{\nu}(v) = 0, InputType'_{\nu}(v) = 1$, this is 1 iff the $\nu$-th predecessor to $v$ is the $i$-th input. This gives $O(N \cdot n)$ variables.
  \item $INPUTVAR_\nu(v,i)$ is equal to $\bigvee_{i'<i} InputVar_\nu(v,i')$, which will be used to bound the fan-in below. This gives $O(N \cdot n)$ variables.
  \item $InputNode_\nu(v,v')$ $(v' < v)$ - when $InputType_{\nu}(v) = 1$, this is 1 iff $\nu$-th predecessor to $v$ is the gate $v'$. This gives $O(N^2)$ variables.
  \item $INPUTNODE_\nu(v,v')$ is analogous to $INPUTVAR_\nu(v,i)$. This gives $O(N^2)$ variables.

@@

  \item Constant predecessors take the correct values, giving $O(N)$ clauses:

@@

  \item Gates have at most one input per input predecessor, giving $O(N \cdot n^2)$ clauses:

@@

  \item INPUTVAR variables take the correct value, giving $O(N \cdot n)$ clauses:

@@

  \item Gates have at least one input per input predecessor, giving $O(N)$ clauses:

@@

  \item Input predecessors take the correct values, giving $O(N \cdot n \cdot 2^n)$ clauses:

@@

  \item Gates take the correct value, giving $O(N \cdot 2^n)$ clauses.

@@

The third reduction is straightforward. It also assumes the function outputs only a single Boolean value and requires a circuit over the basis \(C_2\). The growth rate for the number of clauses is $O(N \cdot 2^n)$, the same as for the Kojevnikov reduction.

@@

Further experiments were conducted on the Oxford ARC service *cite* using 1 node which could run up to 32 processes in parallel. *add more details about ARC pleas

\subsection{Comparison of Boolean functions}\label{funcresults}

We investigated the relationship between the size of a circuit and the time taken for a SAT solver to confirm or deny its existence. We encoded a range of simple symmetric functions using the Kulikov encoding:

\begin{enumerate}
  \item AND on 7-bit and 8-bit inputs
  \item Parity on 3-bit and 4-bit inputs
  \item \(MOD^0_3\) and \(MOD^0_4\) on 3-bit and 4-bit inputs
  \item Majority on 3-bit and 4-bit inputs
  \item Partial parity - *Explain why partial functions are useful for encoding lower bounds on big input-lengths. But also need to find a way to get good data for them (ie. more than 1 second per problem...)

@@

We chose to investigate the existence of circuits up to 9 gates, as these were reliably solved within less than 9 hours on the single-processor hardware described.

The size of the minimal circuit is shown in red*need to re-plot the graphs to make this work better*. It was found that the time needed to solve a problem tended to spike on or around the optimal circuit size. This may be because it is difficult to confirm or rule out circuits that come close to satisfying the requirements. Another possibility is that when close to the optimal size, there is the lowest proportion of satisfying circuits relative to the total amount of possible circuits, meaning that more of the search space must be explored (as opposed to quickly exploring it all at small sizes or quickly landing upon a satisfying circuit at large sizes). A similar phenomenon was reported in~\cite{estrada}.

Different functions took different amounts of time overall, varying between seconds and hours, something also seen in Section \ref{randomresults}. Performance for each function was not especially consistent between the solvers; usually they found the same things difficult, but the length of time each solver took for the difficult problems varied, for example in \ref{fig:8andkulikov} or. The extent of this can be seen in Section \ref{satresults}.

@@

\begin{figure}[!ht]
  \includegraphics[width=\textwidth]{images/times/8andkulikov.png}
  \label{fig:8andkulikov}
  \caption{Times taken by SAT solvers to check for existence of a circuit computing AND on 8 bits}
\end{figure}

\begin{figure}[!ht]
  \includegraphics[width=\textwidth]{images/times/4paritykulikov.png}
  \label{fig:4paritykulikov}
  \caption{Times taken by SAT solvers to check for existence of a circuit computing parity on 4 bits}
\end{figure}

\subsection{Comparison of SAT solvers and reductions} \label{satresults}

We investigated whether there was a relationship between the reduction used to encode a Boolean function, the solver used, and the time taken by the solver. We took the total of the times each combination took to solve the problems from Section \ref{funcresults} and compared them as shown in \ref{totals}.

\begin{figure}[!ht]
  \includegraphics[width=\textwidth]{images/totals.png}
  \label{fig:totals}
  \caption{Comparison of times taken by SAT solvers on different reductions}
\end{figure}

It can be seen that MapleSAT on the Kulikov encoding performs the best. Apart from this, the effect of the reduction and the solver seem to be independent, with MapleSAT consistently performing better than MiniSAT and PicoSAT, and the Naive and Kojevnikov reductions performing better than the Razborov reduction. It is possible that the efficient encoding of the Razborov reduction is detrimental to the performance of the solver. However, these results are only applicable to small values of n. **however I should take out the constants and re-run the experiments to be sure it’s not that..

\subsection{Random functions}\label{randomresults}

64 truth tables were generated for 4-bit inputs using a pseudorandom number generator simulating a uniform distribution (specifically, xorshift128+ due to the TypeScript implementation). The existence of circuits of size 1 to 11 for these tables were encoded with the Kojevnikov reduction and solved using MapleSAT. The times taken to check for circuits are shown plotted against their optimal circuit size in \ref{fig:4rand_scatter} (note that 2 tables could not be fully solved within 9 hours; these functions have been omitted leading to a sample size of 62). The distribution of optimal circuit sizes is shown in \ref{fig:4rand_sizes}.

The scatter graph shows a clustering in times between functions of the same optimal circuit size, but with a few outliers taking more time. This is most easily seen for sizes 10 and 11 (although the same is visible for 8 and 9 when plotted on a smaller scale). Also, the clustered functions of each optimal circuit size took strictly less time to solve than those with larger sizes, despite the fact that we always checked for existence up to 11 gates. This shows a generally strong relationship between the difficulty of checking circuit existence and the complexity of the function.

While we expect a large proportion of circuits to be hard (i.e. have a large optimal circuit size) for large input sizes~\cite{arora}, the usual counting arguments break down for such small values of n. However, using the SAT solver we can check the number of hard circuits directly. While the small sample size can account for the irregular shape of the graph in \ref{fig:4rand_sizes}, there does appear to be a peak around 10 gates. This contradicts the intuition that there should be exponentially more harder functions, and suggests that there is quite a low upper limit on the complexity of a function which only can only vary over 4 bits of input. However, this is all under the assumption that the pseudorandom generator used did output a truly random sample; in reality, it may have sampled disproportionately many easy (or hard) functions.

\begin{figure}[!ht]
  \includegraphics[width=\textwidth]{images/random/4bit_scatter.png}
  \label{fig:4rand_scatter}
  \caption{Time taken to check up to 11 gates for randomly chosen functions *still need to plot on some of the symmetric functions*}
\end{figure}
\begin{figure}[!ht]
  \includegraphics[width=\textwidth]{images/random/4bit_sizes.png}
  \label{fig:4rand_sizes}
  \caption{Sizes of optimal circuits for 4-bit functions with randomly chosen truth tables}

@@

\subsection{Knuth's MOD-3 conjecture}

@@

\subsection{Summary}
section to be included if I need to get closer to 10,000 words :)

@@

The main practical lesson learned from this project was to think in the long term. Firstly, by writing code that can easily be adapted, i.e. following good OOP practices and providing a flexible interface for conducting experiments from the start, so that less refactoring will be needed later. Also, by writing code that can scale to dealing with larger numbers as the scope and ambition of the project expands. This could be done by choosing a well-suited programming language, algorithms and data structures - I chose TypeScript and used arrays as these were familiar and easy to get started with, and the performance of my code and poor documentation of the different pseudorandom generators in the language was an unexpected barrier later on.

describe the general timeline of the project and the different things I implemented that didn't make it into the final, mistakes I made with initial experiments, etc. towards the end would have been better to envision the final experiments and how to get the best results rather than focusing immediately on getting things to work.

things to mention that are not practical - more of how to read papers and do background reading, about how SAT solvers work and their applications, about the state of circuit bounding and P vs NP, the creative process of working on a research project with a lot of backtracking, etc.

@@

\subsection{Further directions}
-Search for the circuit size for literally all 4-bt funcitons (or bigger eg. maybe 9 bits will allow the bounding trick to apply)
